Binning Strategies and Related Loss for Large Data

### 2015-04-10 Paper.Rnw
rescale y axis on origin offset plot
edits on freq loss plot and it's description writeup

<<OriginOffsetSimulations, echo=F,include=T,eval=T,fig.width=8, fig.height=5, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Net spatial loss for coarse simulated data at a resolution of 2 units at various sized (2,4,6,8,10) square bins over the range of possible origin offsets. The vertical gray line in each facet indicates where the origin is offset 1 unit below the minimum data value.">>= 
##  Check out the spatial loss for the data that is recorded to resolution = alpha at various origin offsets
# standard and random binning with bin sizes that are scalar multiples of the data resolution
# save spatial loss information
# spatlossdatRound <- data.frame(NULL)
# index=0
# distindex = 0
# distrib <- c("Uniform", "Normal", "Exponential")
# widths <- seq(alpha,5*alpha,by=alpha)
# for (dat in 1:3){
#   distindex <- distindex + 1
#   for (width in widths){
#     for (origin in seq(-width,0, by=width/60)){
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="standard")
#       binout[[2]]$bintype <- "standard"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="random")
#       binout[[2]]$bintype <- "random"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       print(c(distrib[distindex],width,origin))
#     }
#   }
# }
# write.csv(spatlossdatRound,"LossesByOriginBinwidth.csv", row.names=FALSE)
lossOrigin <- read.csv("LossesByOriginBinwidth.csv",header=TRUE)

qplot(originx,totalSpatialLoss,geom="path",group=distrib, colour=distrib,
      data=subset(lossOrigin,bintype=="standard"), size=I(1))+ 
  facet_grid(distrib~widthx, scales="free_x") + theme_bw() + 
  theme(panel.grid= element_blank(), legend.position="none")+
  geom_vline(xintercept = -1, colour=I("darkgray")) +
  ylab("Net Spatial Loss") + xlab("Origin Offset")
@

@@

The reduced binned data from spatial binning contains the center and count information for all bins.  The bin frequencies may themselves be binned in order to discretize the color scale for the binned scatterplot. The loss of frequency information in frequency binning is dependent on the selection of a discrete color mapping, where the binning algorithm and number of frequency bins must be specified. To explore frequency loss characteristics using the simulated bivariate data sets, each of the three fine resolution sets were first spatially binned with standard rectangular binning with 5 X 5 unit bins and origin at (0,0). Then standard and quantile frequency binning algorithms were applied to either the bin counts or log bin counts. \ktm{ Figure~\ref{freqbinning} shows the binned scatterplots resulting from each type of frequency using five frequency bins.} As discussed in Section~\ref{FreqLoss}, the average person is able to distinguish about seven colors simultaneously, so we will limit the considered options to using between one and ten frequency bins, near the upper limit of human perceptual abilities\citep{colorperception}.

@@

Due to the difference in scales between counts and log counts, the frequency loss and log frequency loss can not be directly compared. The bottom row of Figure~\ref{fig:FreqLossCompare} displays the log frequency loss from using standard and quantile algorithms for binning the \textit{log} bin counts from the same sets of simulated data. Frequency binning using the log counts behaved very similarly, where log frequency loss decreased as the number of frequency bins increased. The primary difference uniform simulated data with bin counts that have low variability, are rendered virtually indistinguishable after a log transformation; thus changing the number of frequency bins does little to impact the log frequency loss. The same advise to use between four and seven frequency bins also holds for shading a binned scatterplot based on log counts.

\begin{figure}[hbtp]
  \begin{minipage}{.3\linewidth}
<<noFreqBin, echo=FALSE, out.width='\\linewidth',fig.width=4, fig.height=4>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
            main="No Freq Binning") +
  xlab("") + ylab("")+
  scale_fill_gradient(low="#56B1F7", high="#132B43") + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     panel.grid= element_blank())
@
  \end{minipage}
\hfil\begin{minipage}{.695\linewidth}
<<FreqBinPlots, echo=FALSE, out.width='\\linewidth',fig.width=10, fig.height=10>>=
binoutFreqGroups <- freqBin(binout1, binType="standard", ncolor=5, logCount=FALSE)
binoutQuantFreqGroups <- freqBin(binout1, binType="quantile", ncolor=5, logCount=FALSE)
binoutLogFreqGroups <- freqBin(binout1, binType="standard", ncolor=5, logCount=TRUE)
binoutQuantLogFreqGroups <- freqBin(binout1, binType="quantile", ncolor=5, logCount=TRUE)

# Plot with Binned Frequencies
labels = paste(unique(binoutFreqGroups[[1]]$freqlabel), "\n(", unique(binoutFreqGroups[[1]]$freqgroup), ")", sep="")
p2 <- qplot(binxs, binys, geom="tile", fill=freqgroup, data=binoutFreqGroups[[1]],
            main="Standard Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_gradient("Frequencies", low="#56B1F7", high="#132B43", 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      breaks=unique(binoutFreqGroups[[1]]$freqgroup),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank())

# Plot with Quantile Binned Frequencies
labels = paste(unique(binoutQuantFreqGroups[[1]]$freqlabel), "\n(", unique(binoutQuantFreqGroups[[1]]$freqgroup), ")", sep="")
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)
p3 <- qplot(binxs, binys, geom="tile", fill=factor(freqgroup), data=binoutQuantFreqGroups[[1]],
            main="Quantile Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_manual("Frequencies", values=cols, 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank()) 
# Plot with standard binned Log Frequencies
labels = paste(unique(binoutLogFreqGroups[[1]]$freqlabel), "\n(", unique(binoutLogFreqGroups[[1]]$freqgroup), ")", sep="")
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)
p4 <- qplot(binxs, binys, geom="tile", fill=freqgroup, data=binoutLogFreqGroups[[1]],
            main="Standard Log Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_gradient("Log Frequencies", low="#56B1F7", high="#132B43", 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      breaks=unique(binoutLogFreqGroups[[1]]$freqgroup),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank())

# Plot with quantile binned Log Frequencies
labels = paste(unique(binoutQuantLogFreqGroups[[1]]$freqlabel), "\n(", unique(binoutQuantLogFreqGroups[[1]]$freqgroup), ")", sep="")
p5 <- qplot(binxs, binys, geom="tile", fill=freqgroup,
            data=binoutQuantLogFreqGroups[[1]],
            main="Quantile Log Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_gradient("Log Frequencies", low="#56B1F7", high="#132B43", 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      breaks=unique(binoutQuantLogFreqGroups[[1]]$freqgroup),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank())

grid.arrange(p2,p3,p4,p5, nrow=2)
@
\end{minipage}
\caption{\label{freqbinning}Simulated bivariate normal data spatially binned using standard algorithm with 5X5 unit bins and origin at (0,0). Binned Scatterplot with raw frequencies (left) and binned scatterplots with five frequency bins (right)."}
\end{figure}

<<FreqLossCompare,echo=F,include=T,eval=T,fig.width=12, fig.height=8, out.width='.99\\linewidth', fig.pos='H',fig.align='center',tidy=F, cache=TRUE, warning=F, fig.show='hold',fig.cap="Lineplots for total frequency loss (top row) and total log frequency loss (bottom row) from standard and quantile binning of bin counts and log bin counts, respectively, using reduced binned data from each bivariate distribution. ">>= 
### Loop over ncolors 1 to 10 and also over data distribution
# find freq loss from standard, quantile, logCount
freqLossSimDat <- NULL
distrib <- c("Uniform", "Normal", "Exponential")
for (distidx in 1:3){
  binout1 <- RectBin2d(datlist[[distidx]]$x,datlist[[distidx]]$y,0,0,5,5,type="standard")
  for (ncol in 1:10){
    binoutFreq <- rbind(freqBin(binout1, binType="standard", ncolor=ncol, logCount=FALSE)[[2]],
                        freqBin(binout1, binType="quantile", ncolor=ncol, logCount=FALSE)[[2]], 
                        freqBin(binout1, binType="standard", ncolor=ncol, logCount=TRUE)[[2]], 
                        freqBin(binout1, binType="quantile", ncolor=ncol, logCount=TRUE)[[2]])
    binoutFreq$FreqBinType <- c("Standard", "Quantile", "Standard Log", "Quantile Log")
    binoutFreq$DistType <- distrib[distidx]
    binoutFreq$ncolor <- ncol 
    freqLossSimDat <- rbind(freqLossSimDat,binoutFreq)
  }
}

@@

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Exponential"),]
p1 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom")

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Normal"),]
p2 <-  qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom")

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Uniform"),]
p3 <-qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom")

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log")  & freqLossSimDat$DistType=="Exponential"),]
p4 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color=FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom")

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log") & freqLossSimDat$DistType=="Normal"),]
p5 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom")

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log") & freqLossSimDat$DistType=="Uniform"),]
p6 <-qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,40) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom")

>>>>

<<OriginOffsetSimulations, echo=F,include=T,eval=T,fig.width=8, fig.height=4, out.width='.9\\linewidth', fig.pos='h',fig.align='center',tidy=F, cache=TRUE, fig.show='hold',fig.cap="Net spatial loss for coarse simulated data at a resolution of 2 units at various sized (2,4,6,8,10) square bins over the range of possible origin offsets. The vertical gray line in each facet indicates where the origin is offset 1 unit below the minimum data value.">>= 
##  Check out the spatial loss for the data that is recorded to resolution = alpha at various origin offsets
# standard and random binning with bin sizes that are scalar multiples of the data resolution
# save spatial loss information
# spatlossdatRound <- data.frame(NULL)
# index=0
# distindex = 0
# distrib <- c("Uniform", "Normal", "Exponential")
# widths <- seq(alpha,5*alpha,by=alpha)
# for (dat in 1:3){
#   distindex <- distindex + 1
#   for (width in widths){
#     for (origin in seq(-width,0, by=width/60)){
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="standard")
#       binout[[2]]$bintype <- "standard"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       index <- index+1
#       binout <- RectBin2d(datlist[[dat]]$roundx,datlist[[dat]]$roundy,origin,origin,width,width,type="random")
#       binout[[2]]$bintype <- "random"
#       binout[[2]]$distrib <- distrib[distindex]
#       spatlossdatRound <- rbind(spatlossdatRound, binout[[2]])
#       print(c(distrib[distindex],width,origin))
#     }
#   }
# }
# write.csv(spatlossdatRound,"LossesByOriginBinwidth.csv", row.names=FALSE)
lossOrigin <- read.csv("LossesByOriginBinwidth.csv",header=TRUE)
qplot(originx,totalSpatialLoss/1000,geom="path",group=distrib, colour=distrib,
      data=subset(lossOrigin,bintype=="standard"), size=I(1))+ 
  facet_grid(distrib~widthx, scales="free_x") + theme_bw() + 
  theme(panel.grid= element_blank(), legend.position="none")+
  geom_vline(xintercept = -1, colour=I("darkgray")) +
  ylab("Net Spatial Loss (in thousands of units)") + xlab("Origin Offset")
@

@@

The reduced binned data from spatial binning contains the center and count information for all bins.  The bin frequencies may themselves be binned in order to discretize the color scale for the binned scatterplot. The loss of frequency information in frequency binning is dependent on the selection of a discrete color mapping, where the binning algorithm and number of frequency bins must be specified. To explore frequency loss characteristics using the simulated bivariate data sets, each of the three fine resolution sets were first spatially binned with standard rectangular binning with 5 X 5 unit bins and origin at (0,0). Then standard and quantile frequency binning algorithms were applied to either the bin counts or log bin counts. As discussed in Section~\ref{FreqLoss}, the average person is able to distinguish about seven colors simultaneously, so we will limit the considered options to using between one and ten frequency bins, near the upper limit of human perceptual abilities\citep{colorperception}.

\ktm{ Figure~\ref{freqbinning} shows the binned scatterplots resulting from standard and quantile frequency binning algorithms using five frequency bins. Note that for standard binning, the shade is to be interpreted as an ordinal indicator of the groups based on bin frequencies, whereas for quantile binning the shade denotes groups based on frequency quantiles.  This distinction explains why the shading of the binned scatterplot is invariant to the log transformation for quantile frequency binning but not for standard; because a monotone transformation will not alter groupings based on order, but will change groupings based on value magnitude.}

@@

Due to the difference in scales between counts and log counts, the frequency loss and log frequency loss can not be directly compared. The bottom row of Figure~\ref{fig:FreqLossCompare} displays the log frequency loss from using standard and quantile algorithms for binning the \textit{log} bin counts from the same sets of simulated data. Frequency binning using the log counts behaved very similarly, where log frequency loss decreased as the number of frequency bins increased. The same advice to use between four and seven frequency bins also holds for shading a binned scatterplot based on log counts.

\begin{figure}[hbtp]
  \begin{minipage}{.31\linewidth}
<<noFreqBin, echo=FALSE, out.width='\\linewidth',fig.width=4, fig.height=4>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
            main="No Freq Binning") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43",
                      guide = guide_colourbar(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top")) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     panel.grid= element_blank(),
                     plot.margin = unit(c(.1,.1,.1,.1),"cm") )
@
  \end{minipage}
\hfil\begin{minipage}{.685\linewidth}
<<FreqBinPlots, echo=FALSE, out.width='\\linewidth',fig.width=10, fig.height=10>>=
binoutFreqGroups <- freqBin(binout1, binType="standard", ncolor=5, logCount=FALSE)
binoutQuantFreqGroups <- freqBin(binout1, binType="quantile", ncolor=5, logCount=FALSE)
binoutLogFreqGroups <- freqBin(binout1, binType="standard", ncolor=5, logCount=TRUE)
binoutQuantLogFreqGroups <- freqBin(binout1, binType="quantile", ncolor=5, logCount=TRUE)

# Plot with Binned Frequencies
labels = paste(unique(binoutFreqGroups[[1]]$freqlabel), "\n(", unique(binoutFreqGroups[[1]]$freqgroup), ")", sep="")
p2 <- qplot(binxs, binys, geom="tile", fill=freqgroup, data=binoutFreqGroups[[1]],
            main="Standard Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_gradient("Frequencies", low="#56B1F7", high="#132B43", 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      breaks=unique(binoutFreqGroups[[1]]$freqgroup),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm"),
                     legend.text=element_text(size=12) )

# Plot with Quantile Binned Frequencies
labels = paste(unique(binoutQuantFreqGroups[[1]]$freqlabel), "\n(", unique(binoutQuantFreqGroups[[1]]$freqgroup), ")", sep="")
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)
p3 <- qplot(binxs, binys, geom="tile", fill=factor(freqgroup), data=binoutQuantFreqGroups[[1]],
            main="Quantile Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_manual("Frequencies", values=cols, 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm"),
                     legend.text=element_text(size=12) ) 
# Plot with standard binned Log Frequencies
labels = paste(unique(binoutLogFreqGroups[[1]]$freqlabel), "\n(", unique(binoutLogFreqGroups[[1]]$freqgroup), ")", sep="")
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)
p4 <- qplot(binxs, binys, geom="tile", fill=freqgroup, data=binoutLogFreqGroups[[1]],
            main="Standard Log Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_gradient("Log Frequencies", low="#56B1F7", high="#132B43", 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      breaks=unique(binoutLogFreqGroups[[1]]$freqgroup),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm") ,
                     legend.text=element_text(size=12))

# Plot with quantile binned Log Frequencies
labels = paste(unique(binoutQuantLogFreqGroups[[1]]$freqlabel), "\n(", unique(binoutQuantLogFreqGroups[[1]]$freqgroup), ")", sep="")
p5 <- qplot(binxs, binys, geom="tile", fill=freqgroup,
            data=binoutQuantLogFreqGroups[[1]],
            main="Quantile Log Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_gradient("Log Frequencies", low="#56B1F7", high="#132B43", 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      breaks=unique(binoutQuantLogFreqGroups[[1]]$freqgroup),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm"),
                     legend.text=element_text(size=12))

grid.arrange(p2,p3,p4,p5, nrow=2)
@
\end{minipage}
\caption{\label{freqbinning}Simulated bivariate normal data spatially binned using standard algorithm with 5X5 unit bins and origin at (0,0). Binned Scatterplot with raw frequencies (left) and binned scatterplots with five frequency bins (right). Bin shades labeled by frequency range and average. }
\end{figure}

<<FreqLossCompare,echo=F,include=T,eval=T,fig.width=12, fig.height=7, out.width='.99\\linewidth', fig.pos='H',fig.align='center',tidy=F, cache=TRUE, warning=F, fig.show='hold',fig.cap="Lineplots for total frequency loss (top row) and total log frequency loss (bottom row) from standard and quantile binning of bin counts and log bin counts, respectively, using reduced binned data from each bivariate distribution. ">>= 
### Loop over ncolors 1 to 10 and also over data distribution
# find freq loss from standard, quantile, logCount
freqLossSimDat <- NULL
distrib <- c("Uniform", "Normal", "Exponential")
for (distidx in 1:3){
  binout1 <- RectBin2d(datlist[[distidx]]$x,datlist[[distidx]]$y,0,0,5,5,type="standard")
  for (ncol in 1:10){
    binoutFreq <- rbind(freqBin(binout1, binType="standard", ncolor=ncol, logCount=FALSE)[[2]],
                        freqBin(binout1, binType="quantile", ncolor=ncol, logCount=FALSE)[[2]], 
                        freqBin(binout1, binType="standard", ncolor=ncol, logCount=TRUE)[[2]], 
                        freqBin(binout1, binType="quantile", ncolor=ncol, logCount=TRUE)[[2]])
    binoutFreq$FreqBinType <- c("Standard", "Quantile", "Standard Log", "Quantile Log")
    binoutFreq$DistType <- distrib[distidx]
    binoutFreq$ncolor <- ncol 
    freqLossSimDat <- rbind(freqLossSimDat,binoutFreq)
  }
}

@@

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Exponential"),]
p1 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Normal"),]
p2 <-  qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Uniform"),]
p3 <-qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log")  & freqLossSimDat$DistType=="Exponential"),]
p4 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color=FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log") & freqLossSimDat$DistType=="Normal"),]
p5 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log") & freqLossSimDat$DistType=="Uniform"),]
p6 <-qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,2) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))
